M.Sc. Proposal: \\ \bigskip Computational Electrokinetics

### 2013-01-03 notes/thesis-latex-template/thesis.tex
Update thesis.

% Bold symbols and operators
\newcommand\Laplacian{\nabla^2}
\newcommand\bnabla{\boldsymbol{\nabla}}
\newcommand\bLaplacian{\boldsymbol{\nabla}^2}
\newcommand\bcdot{\boldsymbol{\cdot}}
\newcommand\bU{\mathscr{\boldsymbol{U}}}
\newcommand\bv{\boldsymbol{v}}
\newcommand\bV{\boldsymbol{V}}
\newcommand\bE{\boldsymbol{E}}
\newcommand\be{\boldsymbol{\hat{e}}}
\newcommand\bn{\boldsymbol{\hat{n}}}
\newcommand\bk{\boldsymbol{\hat{k}}}
\newcommand\bj{\boldsymbol{j}}
\newcommand\bi{\boldsymbol{i}}
\newcommand\bs{\boldsymbol{s}}
\newcommand\bA{\boldsymbol{A}}
\newcommand\bF{\boldsymbol{F}}
\newcommand\bG{\boldsymbol{G}}
\newcommand\bI{\boldsymbol{I}}
\newcommand\bJ{\boldsymbol{J}}
\newcommand\bx{\boldsymbol{x}}
\newcommand\by{\boldsymbol{y}}
\newcommand\bz{\boldsymbol{z}}
\newcommand\br{\boldsymbol{r}}
\newcommand\bc{\boldsymbol{c}}
\newcommand\bxhat{\hat{\bx}}
\newcommand\byhat{\hat{\by}}
\newcommand\bzhat{\hat{\bz}}
\newcommand\brhat{\hat{\br}}
\newcommand\bnhat{\hat{\boldsymbol{n}}}
\newcommand\btheta{\boldsymbol{\theta}}
\newcommand\bthetahat{\hat{\btheta}}
\newcommand\bphi{\boldsymbol{\phi}}
\newcommand\bphihat{\hat{\bphi}}
\newcommand\bzero{\boldsymbol{0}}
\newcommand\bomega{\boldsymbol{\omega}}
\newcommand\bpsi{\boldsymbol{\psi}}

@@

\subsubsection{Highly charged inert particle}
See \cite{schnitzer2012surface} for full derivation
of the nonlinear effective boundary conditions. 
Note that due to surface conduction, the resulting boundary conditions are:
\begin{align} \label{eq:ephor_bnd}
\nonumber
0 &= {\deriv{C}{n} - \text{Du} \nabla^2_s \pars{\varPhi - \log C}}, \\
0 &= {\deriv{C}{n} + C \deriv{\varPhi}{n}}, \\
\nonumber
\bV &= \zeta \cdot \bnabla_\mathcal{S} \pars{\varPhi - \log C} +
4 \log 2 \cdot \bnabla_\mathcal{S} \log C,
\end{align}
where $\zeta = \bar{\zeta} - \log C$ 
is the zeta potential and $\bar{\zeta} = 2 \log \sigma$.

@@

\subsubsection{Highly charged inert particle} The boundary conditions \eqref{eq:ephor_bnd} 
are implemented by:
\begin{equation}
\begin{array}{rcl}
0 &=& \cD_r(C^h) - \text{Du} \cdot 
\pars{\cD_\theta \pars{\sin\theta \cdot \cD_\theta \cI_r\pars{\varPhi^h - \log C}}}/{\sin\theta}, \\
0 &=& \cD_r(C^h) + \cI_r(C^h) \cD_r(\varPhi^h), \\
V^h_r &=& 0, \\
V^h_\theta &=& \cI_\theta(\zeta^h) \cdot \cD_\theta(\cI_r \varPhi^h - \log \cI_r C^h) + 
4 \log 2 \cdot \cD_\theta(\log \cI_r C^h), \\
  \zeta^h &=& \bar{\zeta} - \log \cI_r C.
\end{array}
\end{equation}

@@

For $\beta = 0$ (no electric field is applied), the steady-state solution is $\cU = 0$:
\begin{equation}\begin{array}{cccc}
\varPhi_0(r,\theta) = 0, &
C_0(r,\theta) = 1, &
\bV_0(r,\theta) = \bzero, &
P_0(r,\theta) = 0.
\end{array}\end{equation}
Assuming $\beta \ll 1$, the linearized equations and boundary conditions are
\begin{equation} \begin{array}{ccc}
\Laplacian \varPhi_1 = 0, &
\Laplacian C_1 = 0, &
\bLaplacian \bV_1 - \bnabla P_1 = \bzero.
\end{array}\end{equation}
The first-order terms $\bx_1$ are given by (see \cite{yariv2010migration} for full derivation)
\begin{equation} \begin{array}{cccc}
\varPhi_1 = \pars{\frac{1}{4r^2} - r}\cos\theta, &
C_1 = \frac{3}{4r^2} \cos\theta, &
\Psi_1 = \cU_1 \pars{\frac{1}{r} - r^2} \frac{\sin^2\theta}{2}, &
P_1 = 0,
\end{array} \end{equation}
and the linear velocity term is:
\begin{align} \label{eq:linear_velocity}
\bV_1 &= \cU_1 \pars{ \pars{-1 + r^{-3}}\cos\theta \cdot \brhat 
+ \pars{1 + r^{-3} / 2} \sin\theta \cdot \bthetahat } \\
\cU_1(\gamma) &= 2 \log \pars{\frac{1 + \gamma^{-\frac{1}{2}}}{2}}.
\end{align}
For $\gamma > 1$, the Debye layer has a positive space charge, 
corresponding to a negative particle charge -- 
yielding negative drift velocity, $\cU < 0$. 
For $\gamma < 1$, we have $\cU > 0$.

@@

Since the leading order of the solution $\cU(\beta)$ is linear for $\beta \ll 1$, 
the numerical discretization error is $O(\beta)$ as well.
In order to reduce the discretization error of the numerical solution \eqref{eq:disc_zero_force}
and after verifying that the discretization error converges as $O(h^2)$,
we employ Richardson extrapolation to $\cU^h$ and $\cU^{2h}$:
\begin{equation}
\hat\cU = \frac{4}{3} \cU^h - \frac{1}{3} \cU^{2h}.
\end{equation}
The results of the extrapolation and $\hat\cU$ the analytical results 
are given at Figures \ref{fig:Electrophoresis1} and \ref{fig:Electrophoresis2}.
The numerical solver uses $R_{\max} = 10^2$,
$K = 3$, $M = 5$ and $\beta = 0.1 \cdot 2^{t}$ for $t = -2:0.5:5$.
The solver is applied to 2 grids of $N_r \times N_\theta = (128 \times 128)$ as the coarse grid 
(with $\Delta_r \approx 0.037$), and $N_r \times N_\theta = (256 \times 256)$ as the fine grid
(with $\Delta_r \approx 0.018$), in order to use Richardson extrapolation for $\hat\cU$.

@@

The computational grid is created to have fine resolution near the particle surface, while
approximating the far-away boundary conditions. The logarithmic discretization is motivated
by the solution of the linearized Stokes problem, where the largest gradients are near the particle
and the solution approaches the outer boundary conditions far away from the particle.

@@

>>>>

% Bold symbols and operators
\newcommand\Laplacian{\nabla^2}
\newcommand\bnabla{\boldsymbol{\nabla}}
\newcommand\bLaplacian{\boldsymbol{\nabla}^2}
\newcommand\bcdot{\boldsymbol{\cdot}}
\newcommand\bU{\mathscr{\boldsymbol{U}}}
\newcommand\bv{\boldsymbol{v}}
\newcommand\bV{\boldsymbol{V}}
\newcommand\bE{\boldsymbol{E}}
\newcommand\be{\boldsymbol{\hat{e}}}
\newcommand\bn{\boldsymbol{\hat{n}}}
\newcommand\bk{\boldsymbol{\hat{k}}}
\newcommand\bj{\boldsymbol{j}}
\newcommand\bi{\boldsymbol{i}}
\newcommand\bs{\boldsymbol{s}}
\newcommand\bA{\boldsymbol{A}}
\newcommand\bF{\boldsymbol{F}}
\newcommand\bG{\boldsymbol{G}}
\newcommand\bI{\boldsymbol{I}}
\newcommand\bJ{\boldsymbol{J}}
\newcommand\bx{\boldsymbol{x}}
\newcommand\by{\boldsymbol{y}}
\newcommand\bz{\boldsymbol{z}}
\newcommand\br{\boldsymbol{r}}
\newcommand\bc{\boldsymbol{c}}
\newcommand\bxhat{\hat{\bx}}
\newcommand\byhat{\hat{\by}}
\newcommand\bzhat{\hat{\bz}}
\newcommand\brhat{\hat{\br}}
\newcommand\bnhat{\hat{\boldsymbol{n}}}
\newcommand\btheta{\boldsymbol{\theta}}
\newcommand\bthetahat{\hat{\btheta}}
\newcommand\bphi{\boldsymbol{\phi}}
\newcommand\bphihat{\hat{\bphi}}
\newcommand\bzero{\boldsymbol{0}}
\newcommand\bomega{\boldsymbol{\omega}}
\newcommand\bpsi{\boldsymbol{\psi}}
\newcommand\Du{\text{Du}}

@@

\subsubsection{Highly charged inert particle}
See \cite{schnitzer2012surface} for full derivation
of the nonlinear effective boundary conditions. 
Note that due to surface conduction, the resulting boundary conditions are:
\begin{align} \label{eq:ephor_bnd}
\nonumber
0 &= {\deriv{C}{n} - \Du \nabla^2_s \pars{\varPhi - \log C}}, \\
0 &= {\deriv{C}{n} + C \deriv{\varPhi}{n}}, \\
\nonumber
\bV &= \zeta \cdot \bnabla_\mathcal{S} \pars{\varPhi - \log C} +
4 \log 2 \cdot \bnabla_\mathcal{S} \log C,
\end{align}
where $\zeta = \bar{\zeta} - \log C$ 
is the zeta potential and $\bar{\zeta} = 2 \log \sigma$.

@@

\subsubsection{Highly charged inert particle} The boundary conditions \eqref{eq:ephor_bnd} 
are implemented by:
\begin{equation}
\begin{array}{rcl}
0 &=& \cD_r(C^h) - \Du \cdot 
\pars{\cD_\theta \pars{\sin\theta \cdot \cD_\theta \cI_r\pars{\varPhi^h - \log C}}}/{\sin\theta}, \\
0 &=& \cD_r(C^h) + \cI_r(C^h) \cD_r(\varPhi^h), \\
V^h_r &=& 0, \\
V^h_\theta &=& \cI_\theta(\zeta^h) \cdot \cD_\theta(\cI_r \varPhi^h - \log \cI_r C^h) + 
4 \log 2 \cdot \cD_\theta(\log \cI_r C^h), \\
  \zeta^h &=& \bar{\zeta} - \log \cI_r C.
\end{array}
\end{equation}

@@

For $\beta = 0$ (no electric field is applied), the steady-state solution is $\cU = 0$:
\begin{equation}\begin{array}{cccc}
\varPhi_0(r,\theta) = 0, &
C_0(r,\theta) = 1, &
\bV_0(r,\theta) = \bzero, &
P_0(r,\theta) = 0.
\end{array}\end{equation}
Assuming $\beta \ll 1$, the linearized equations and boundary conditions are
\begin{equation} \begin{array}{ccc}
\Laplacian \varPhi_1 = 0, &
\Laplacian C_1 = 0, &
\bLaplacian \bV_1 - \bnabla P_1 = \bzero.
\end{array}\end{equation}
The first-order terms $\bx_1$ are given by (see \cite{yariv2010migration} for full derivation)
\begin{equation} \begin{array}{cccc}
\varPhi_1 = \pars{\frac{1}{4r^2} - r}\cos\theta, &
C_1 = \frac{3}{4r^2} \cos\theta, &
\Psi_1 = \cU_1 \pars{\frac{1}{r} - r^2} \frac{\sin^2\theta}{2}, &
P_1 = 0,
\end{array} \end{equation}
and the linear velocity term is:
\begin{align} \label{eq:linear_velocity}
\bV_1 &= \cU_1 \pars{ \brhat \pars{-1 + \frac{1}{r^3}}\cos\theta 
+ \bthetahat \pars{1 + \frac{1}{2r^3}} \sin\theta } \\
\cU_1(\gamma) &= 2 \log \pars{\frac{1 + \gamma^{-\frac{1}{2}}}{2}}.
\end{align}
For $\gamma > 1$, the Debye layer has a positive space charge, 
corresponding to a negative particle charge -- 
yielding negative drift velocity, $\cU < 0$. 
For $\gamma < 1$, we have $\cU > 0$.

@@

Since the leading order of the solution $\cU(\beta)$ is linear for $\beta \ll 1$, 
the numerical discretization error is $O(\beta)$ as well.
In order to reduce the discretization error of the numerical solution \eqref{eq:disc_zero_force}
and after verifying that the discretization error converges as $O(h^2)$ (Figure \ref{fig:quadratic}),
we employ Richardson extrapolation to $\cU^h$ and $\cU^{2h}$:
\begin{equation}
\hat\cU = \frac{4}{3} \cU^h - \frac{1}{3} \cU^{2h}.
\end{equation}
The results of the extrapolation and $\hat\cU$ the analytical results 
are given at Figures \ref{fig:Electrophoresis1} and \ref{fig:Electrophoresis2}.
The numerical solver uses $R_{\max} = 10^2$,
$K = 3$, $M = 5$ and $\beta = 0.1 \cdot 2^{t}$ for $t = -2:0.5:5$.
The solver is applied to 2 grids of $N_r \times N_\theta = (128 \times 128)$ as the coarse grid 
(with $\Delta_r \approx 0.037$), and $N_r \times N_\theta = (256 \times 256)$ as the fine grid
(with $\Delta_r \approx 0.018$), in order to use Richardson extrapolation for $\hat\cU$.

@@

The computational grid is created to have fine resolution near the particle surface, while
approximating the far-away boundary conditions. The logarithmic discretization is motivated
by the solution of the linearized system at Section \ref{app:linear}, 
where the largest gradients in the solution are near the particle surface.
The discretization of boundary conditions far away from the particle may be improved
by including the linear correction from the asymptotic solution, or applying

@@

\begin{figure}
    \begin{center}
    \includegraphics[width=1\textwidth]{figs/quadratic_order.eps}
        \caption[Quadratic convergence of steady-state velocity]{
        The discretization error between the numerical and analytical solutions for 
        $\alpha = 0$ ,$\Du = 0.5$, $\bar\zeta = 6$ and $R_{max} = 100$. The markers represent
        the discretization errors for various grid sizes, and the dotted lines correspond to linear
        functions of $\beta$, scaled by $4^n$. As mesh cell size decreases by 2, the discretization
        }
	    \label{fig:quadratic}
    \end{center}
\end{figure}
